---
title: "R language cheatsheet"
author: "Geo. Naumov"
date: "2024-11-15"
output: pdf_document
---

# R language cheatsheet

## Basic calculator
```{r}
10 / 2
11 / 2.1
6 * 7
(15 + 12) / ( 19 - 12)
factorial(10)
6 ^ 2
x <- 3
4 -> y
z <- x*y
print(z)
```

## Output
```{r}
print(z)
print("Hello from R")

hello_str <- "Hello from R"
print(hello_str)
```

## Comparing variables
```{r}
result1 <- x<y
result2 <- x>y
result3 <- x==y
```

## Vectorized programming language
```{r}
typeof(z)
length(z)

typeof(hello_str)
length(hello_str)

```

### Combine collections as vectors

DUE TO THIS, A VARIABLE CCANNOT BE CALLED c BECAUSE THAT WILL OVERWRITE THE COLLECTION FUNCTION!!!

```{r}

# Create a vector using a collection 
col1 <- c(3,5,6,7)
typeof(col1)
length(col1)

col2 <- c("Hello", "man")
typeof(col2)
length(col2)

# Access vector items
col2[1]
col1[2]
col1[1:3]
```

### Operation between vectors
Vectors must be of same size!
```{r}
vec1 <- c(3,5,6,7)
vec2 <- c(2,4,8,9)
vec3 <- vec1 * vec2
vec4 <- vec1 + vec2
```
A single item vector can also be applied to a larger vector. 

In this case, it will be appended n-times by the same value in order to become as long as the vector we are operating on. Then all values are calculated as in the previous example.

The example below multiplies everything in vec1 by 2, result is vector of same size as vec1

```{r}
vec5 <- vec1 * 2 
```

## Data types

### Vector data types
  Integers - Numbers
  Doubles - Numbers with a fraction associated with it
  Characters - Text, strings
  Logicals - Booleans true/false
  
###  Non-vector data types
  Lists - Vectors, but with items of different types
  Functions - you know ...
  Dataframes - Tabular data

#### Lists
```{r}
list1 <- list(1,2,"Hello","man",TRUE,FALSE)
typeof(list1)
length(list1)
```
For comparison, a vector with the same data
```{r}
vec6 <- c(1,2,"Hello","man",TRUE,FALSE)
typeof(vec6)
length(vec6)
```
The vector is of type character because vectors can have only one data type, text is the most complex one that can incorporate the rest. 

This does not apply to lists!

R converts data types automatically.

#### List of vectors
R is vectorized, lists are not vectors but the items inside - are!

```{r}
employee_list <- list(name = "Geo", 
                      surname = "Naumov", 
                      age = 36, 
                      statistics = TRUE,
                      courses = c(10,11,12),
                      12345,
                      "bla bla"
                      )

print(employee_list)
```
Here we have a list of vectors with some named values of different types, a vector, and also 2 unnamed items: 12345 and "bla bla".

Items start with $ sign and it can be used to access them.

```{r}
employee_list$name
employee_list$statistics
employee_list$courses
```
Another way is by index:
```{r}
employee_list[1]
employee_list[6]
employee_list[7]
```
This is also the way to access unnamed items.

#### Attributes
Some operations on attributes
```{r}
attributes(employee_list)

# Does not work, it's a separate list created by R
typeof(attributes(employee_list))

# Get the attribute names, as strings in a vector
attribute_names <- names(employee_list)

# Set attribute name
attribute_names[6] <- "some_number"
attribute_names[7] <- "some_text"

# Assign the data back to employee_list because setNames returns a copy
employee_list <- setNames(employee_list, attribute_names)

attributes(employee_list)
```
### Null values
Assign a null value to the variable z:
```{r}
z <- NULL
typeof(z)
length(z)
# z == NULL # Not recommended, you cannot compare nothing to nothing
is.null(z)  # The correct way
```
### NA values
A single item vector with a value of NA
```{r}
x <- NA
typeof(x)
length(x)
is.na(x)
```
