---
title: "02 Dataframes"
author: "Geo. Naumov"
date: "2024-11-15"
output: html_document
---

# Dataframes

## About dataframes
In dataframes, rows have a single data type. In R, columns are called variables and rows are called observations, because R is a statistical language. 

Names of columns are collected from the first row of the CSV file. Rows usually don't have names.

Columns can be accessed using the $ sign and rows using a range in blocks [].

R pays a attention to the order of rows and they are loaded in the exact same order as the original data.

Rows have many different data types between different columns, so they cannot be extracted as vectors but rather as lists.

Columns have the same data type so they can be extracted as vectors.


## Load the data
```{r}
df_admissions <- read.csv("adm2020.csv")
```


## Select some data

### Rows and columns
```{r}
# First row
first_row <- df_admissions[1,]

# First 10 rows, values are inclusive
first_10_row <- df_admissions[1:10,]

# Last 10 columns, values are inclusive
last_10_col <- df_admissions[,58:68]

# First 10 columns and first 10 rows
first_10_col_row <- df_admissions[1:10,58:68]

# Columns can be selected by name
some_cols <- df_admissions[,c("ADMCON1", "ADMCON2", "ADMCON3")]

# Rows by index and columns by names
some_rows_cols <- df_admissions[1:10,c("ADMCON1", "ADMCON2", "ADMCON3")]
```
The comma is called a subsetting operator. 

### Skip data
Use this to select data but skip some columns or rows.
```{r}
# Select rows 1 to 10 but exclude columns 1,3,5
some_skipped_columns <- df_admissions[1:10, c(-1,-3,-5)]

```
